---
title: "signals from all clusters to senescent populations and receptors connected downstream to SRF, TEAD4"
author: "Sushma Nagaraj"
date: "2024-11-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

suppressPackageStartupMessages({
  library(domino2)
  library(purrr)
  library(dplyr)
  library(tidyr)
  library(tibble)
  library(ComplexHeatmap)
  library(circlize)
  library(stringr)
  library(RColorBrewer)
})

here::i_am("3_domino/all_signaling_to_SnC.receptors_connected_to_SRF_TEAD4.Rmd")
source("functions/network2df.R")
```


```{r}
clusters_of_interest <- c("Fibroblast_SnC", "Pericyte_SMC_SnC", "Endo_vein2_SnC")
```


```{r}
stromal <- c("Endo_artery", "Endo_vein1", "Endo_vein2_not_SnC", "Endo_vein2_SnC", "Lymphatic_endo_not_SnC", "Lymphatic_endo_SnC", "Fibroblast_not_SnC", "Fibroblast_SnC", "Pericyte_SMC_not_SnC", "Pericyte_SMC_SnC", "SMC_not_SnC", "SMC_SnC")
immune <- c("Monocytes", "Macrophages", "Myeloid", "DC", "Granulocytes", "T_cells", "NK") 
all <- c(stromal, immune)
```


# Read in the rec_lig differential linkages
```{r}
diff_linkages_list <- lapply(clusters_of_interest, function(cluster) {
  read.csv(here::here("results", "3_domino2", "cell_type_SnC", "differential_linkages_rec_lig_activation",  paste0(cluster, ".differential_linkages.csv")), header = T)
})
names(diff_linkages_list) <- clusters_of_interest
```


# Retain only the linkages found in at least 3 of 4 fibroid samples
```{r}
diff_linkages <- list_rbind(diff_linkages_list)

diff_linkages <- diff_linkages %>%
  filter(Fibroid_count >= 3)

diff_linkages$key <- paste0(diff_linkages$cluster, "_", diff_linkages$feature)
```

# Read in domino object
```{r}
dom <- readRDS(here::here("results", "3_domino2", "cell_type_SnC", "Fibroid.domino.rds"))
```


```{r}
network_all <- network2df(dom = dom, sending_cluster = all, rec_clusters = clusters_of_interest)
network_all$key <- paste0(network_all$receiving_cluster, "_", network_all$receptor, " <- ", network_all$ligand)
```


```{r}
network <- network_all %>%
  filter(key %in% diff_linkages$key)

title <- paste0("All_to_SnC")
```


# Chord plot functions
```{r}
circos_plot <- function(network_slice) {
  set.seed(1)
  chordDiagram(network_slice[,c("receptor", "transcription_factor", "cluster_cor")],
              col = network_slice$col, 
              big.gap = 20,
              #scale = TRUE,
              directional = 1, 
              direction.type = c("diffHeight", "arrows"),
              #direction.type = c("arrows"),
              link.arr.type = "big.arrow", 
              annotationTrack = "grid",
              preAllocateTracks = list(list(track.height = 0.2),
                                       list(track.height = 0.08)))

  circos.track(track.index = 1,
               panel.fun = function(x, y) {
               circos.text(CELL_META$xcenter, 
                           CELL_META$ylim[1], 
                           #str_wrap(CELL_META$sector.index, width = 20), 
                           str_split(CELL_META$sector.index, pattern = "\\.") |> unlist() |> tail(1),
                           facing = "clockwise", 
                           cex = 0.8, 
                           adj = c(0, 0.5), 
                           niceFacing = TRUE)},
               bg.border = NA)
circos.clear()
}
```

# Color scale
```{r}
cluster_cols <- c("Fibroblast_not_SnC", "Pericyte_SMC_not_SnC", "SMC_not_SnC", "Epithelial", "Cycling", "T_cells", "NK", "Lymphatic_endo_not_SnC", "Endo_artery", "Endo_vein2_not_SnC", "Endo_vein1", "DC", "Macrophages", "Monocytes", "Myeloid","Granulocytes", "Fibroblast_SnC", "Pericyte_SMC_SnC", "Lymphatic_endo_SnC", "Endo_vein2_SnC", "SMC_SnC")
names(cluster_cols) <- c("#E69F00", "#56B4E9", "#009E73", "#0072B2", "#D55E00", "#CC79A7", "#7B7B7B", "#FFCB57", "#AADAF4", "#00EEB0", "#F7F19F", "#FFB679", "#E6BCD3", "#954B01", "#8661D1","#BFA4E4", "#b37a00", "#2ba0e3", "#ffba24", "#00bd8b", "#8B820B")
```

# Circos plot for TF-rec of interest
```{r}
tf <- c("SRF", "TEAD4")

cols <- setNames(names(cluster_cols), cluster_cols)
network_slice <- network %>%
  filter(transcription_factor %in% tf) %>%
  mutate(col = recode(receiving_cluster, !!!cols)) %>% 
  select(receptor, transcription_factor, cluster_cor, receiving_cluster, col) %>%
  filter(cluster_cor > 0) %>%
  distinct()

# Cluster legend
network_slice %>% 
  select(receiving_cluster, col) %>%
  distinct %>%
  deframe -> cl_cols
cl_legend = Legend(labels = names(cl_cols), title = "Cluster", legend_gp = gpar(fill = cl_cols))

circos_plot(network_slice)
draw(cl_legend, x = unit(2, "mm"), y = unit(6, "mm"), just = c("left", "bottom"))
```
```{r, echo=FALSE, eval=FALSE}
pdf(here::here("plots_for_figures", "plots", "Fig_5", paste0(title, "_receptors_connected_to_SRF_TEAD4.chord_diagram.pdf")))
circos_plot(network_slice)
draw(cl_legend, x = unit(2, "mm"), y = unit(6, "mm"), just = c("left", "bottom"))
dev.off()

write.csv(network_slice, here::here("plots_for_figures", "data_tables", "Fig_5", "data_in_plots", paste0(title, "_receptors_connected_to_SRF_TEAD4.chord_diagram.csv")))
```


# Heatmaps
# Heatmap plotting function
```{r}
incoming_to_one_cluster <- function(network, title, filter=FALSE, fthresh="Mean") {
  network.wide <- network %>%
    select(sending_cluster, product_rec_lig, rec_lig) %>%
    unique() %>%
    pivot_wider(names_from = sending_cluster, values_from = product_rec_lig, id_cols = rec_lig) %>%
    column_to_rownames(var = "rec_lig")

  network.wide[is.na(network.wide)] <- 0
  if(filter) {
    thresh <- summary(rowSums(network.wide))[[fthresh]]
    network.wide <- network.wide[rowSums(network.wide) > thresh, ]
  } 
  
  ht <- Heatmap(as.matrix(network.wide),
                  cluster_rows = T, 
                  cluster_columns = T,
                  col = brewer.pal(n = 9, name = "YlOrRd"),
                  #clustering_distance_rows = "pearson",
                  #clustering_distance_columns = "pearson",
                  show_column_dend = F,
                  show_row_dend = F,
                  #column_title_rot = 90,
                  name = "rec_exp * ligand_exp",
                  row_title = "Interactions",
                  row_title_side = "left",
                  row_names_gp = gpar(fontsize = 9),
                  column_title = title
  )
  return(ht)
}
```

# Heatmap showing ligands signaling to the receptors associated with TEAD4 and SRF
```{r}
network_slice <- network %>%
  filter(transcription_factor %in% tf) %>%
  filter(cluster_cor > 0.1) %>%
  filter(! sending_cluster %in% c("Lymphatic_endo_not_SnC", "Lymphatic_endo_SnC", "SMC_not_SnC", "SMC_SnC"))

# Additional columns to network df for plotting
network_slice$rec_lig <- paste0(network_slice$receptor, " <- ", network_slice$ligand)
network_slice$product_rec_lig <- network_slice$ligand_mean_z_score * network_slice$receptor_mean_z_score
```

# All to Fibroblast_SnC
```{r, fig.height=8}
fib <- network_slice %>%
  filter(receiving_cluster == "Fibroblast_SnC")
title <- "All to SnC Fibro"
ht <- incoming_to_one_cluster(fib, title)
ht
```


```{r, echo=FALSE, eval=FALSE}
pdf(here::here("plots_for_figures", "plots", "SuppFig_15", paste0(title, "_ligands _to_receptors_connected_to_SRF_TEAD4.cluster_cor_gt0.1.heatmap.pdf")))
ht
dev.off()
```

# All to Pericyte_SMC_SnC
```{r, fig.height=8}
peri <- network_slice %>%
  filter(receiving_cluster == "Pericyte_SMC_SnC")
title <- "All to SnC Mural"
ht <- incoming_to_one_cluster(peri, title)
ht
```


```{r, echo=FALSE, eval=FALSE}
pdf(here::here("plots_for_figures", "plots", "SuppFig_15", paste0(title, "_ligands _to_receptors_connected_to_SRF_TEAD4.cluster_cor_gt0.1.heatmap.pdf")))
ht
dev.off()
```

# All to Endo_vein2_SnC
```{r}
endo <- network_slice %>%
  filter(receiving_cluster == "Endo_vein2_SnC")
title <- "All to SnC EC2"
ht <- incoming_to_one_cluster(endo, title)
ht
```


```{r, echo=FALSE, eval=FALSE}
pdf(here::here("plots_for_figures", "plots", "SuppFig_15", paste0(title, "_ligands _to_receptors_connected_to_SRF_TEAD4.cluster_cor_gt0.1.heatmap.pdf")))
ht
dev.off()
```


```{r}
sessionInfo()
```








