---
title: "SnC clusters ligands communication networks"
author: "Sushma Nagaraj"
date: "2026-05-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

suppressPackageStartupMessages({
  library(domino2)
  library(purrr)
  library(dplyr)
  library(tidyr)
  library(tibble)
  library(forcats)
  library(ComplexHeatmap)
  library(circlize)
  library(stringr)
  library(igraph)
})

here::i_am("3_domino/SnC_vascular_to_stromal_vascular_network.Rmd")
source("functions/network2df.R")
source("functions/dom_df_to_igraph.R")

## Fig_3
```


```{r}
tissue <- "Fibroid"
tissue
```

# Read in the rec_lig differential linkages
```{r}
clusters_of_interest <- c("Endo_artery", "Endo_vein1", "Endo_vein2_not_SnC", "Fibroblast_not_SnC",  "Pericyte_SMC_not_SnC")

diff_linkages_list <- lapply(clusters_of_interest, function(cluster) {
  read.csv(here::here("results", "3_domino2", "cell_type_SnC", "differential_linkages_rec_lig_activation",  paste0(cluster, ".differential_linkages.csv")), header = T)
})
names(diff_linkages_list) <- clusters_of_interest
```

# Retain only the linkages found in at least 3 of 4 fibroid samples/ 3 myometrium samples
```{r}
diff_linkages <- list_rbind(diff_linkages_list)

#Fibroid
diff_linkages <- diff_linkages %>%
  filter(Fibroid_count >= 3)

diff_linkages$key <- paste0(diff_linkages$cluster, "_", diff_linkages$feature)
```

# Network with ligands from SnC Celltypes
```{r}
dom <- readRDS(here::here("results", "3_domino2", "cell_type_SnC", paste0(tissue, ".domino.rds")))
```


```{r}
cluster <- c("Pericyte_SMC_SnC", "Endo_vein2_SnC")
cluster
```

# Extract networks for cluster from the Domino object
```{r}
outgoing_network <- network2df(dom = dom, sending_cluster = cluster, rec_clusters = clusters_of_interest)
```


```{r}
outgoing_network$key <- paste0(outgoing_network$receiving_cluster, "_", outgoing_network$receptor, " <- ", outgoing_network$ligand)
outgoing_network$ligand_key <- paste0(outgoing_network$sending_cluster, "_", outgoing_network$ligand)
```

# Read in differentially expressed secreted/Non-secreted for cluster from SnC-vs-nonSnC analysis
```{r}
cl <- gsub("_SnC", "", cluster)

de_genes <- lapply(cl, function(x) {
  SnC.genes <- read.csv(here::here("results", "2_SnC-vs-nonSnC", str_to_lower(tissue), paste0(str_to_lower(tissue), ".", x, ".SnC.genes.csv")), header = TRUE)

  DE_subset <- SnC.genes %>%
    filter(p_val_adj < 0.05 & abs(avg_log2FC) > 0.3)
})
names(de_genes) <- cl

de_genes <- list_rbind(de_genes, names_to = "cell_type")
de_genes$cell_type <- paste0(de_genes$cell_type, "_SnC")
de_genes$ligand_key <- paste0(de_genes$cell_type, "_", de_genes$X)
```

# Networks for ligands present in atleast 3 fibroid samples and DE in SnC populations (compared to non-SnC)
```{r}
# Colors for network plots
cluster_cols <- c("Fibroblast_not_SnC", "Pericyte_SMC_not_SnC", "SMC_not_SnC", "Epithelial", "Cycling", "T_cells", "NK", "Lymphatic_endo_not_SnC", "Endo_artery", "Endo_vein2_not_SnC", "Endo_vein1", "DC", "Macrophages", "Monocytes", "Myeloid","Granulocytes", "Fibroblast_SnC", "Pericyte_SMC_SnC", "Lymphatic_endo_SnC", "Endo_vein2_SnC", "SMC_SnC")
names(cluster_cols) <- c("#E69F00", "#56B4E9", "#009E73", "#0072B2", "#D55E00", "#CC79A7", "#7B7B7B", "#FFCB57", "#AADAF4", "#00EEB0", "#F7F19F", "#FFB679", "#E6BCD3", "#954B01", "#8661D1","#BFA4E4", "#b37a00", "#2ba0e3", "#ffba24", "#00bd8b", "#8B820B")

```

# Signaling to Stromal
```{r}
out.net.sub.stromal <- outgoing_network[(outgoing_network$ligand_key %in% de_genes$ligand_key) & (outgoing_network$key %in% diff_linkages$key), ]

out.net.sub.stromal <- out.net.sub.stromal %>%
  filter(!receptor == "integrin_a1b1_complex")
```

# Nazmiye's curation - filter interactions without literature support
```{r}
out.net.sub.stromal <- out.net.sub.stromal %>%
  filter(!ligand %in% c("SCGB3A1") & !receptor %in% c("AR"))
```

## Network plot
```{r, fig.height=8}
send_cols <- cluster_cols[cluster_cols %in% cluster]
rec_cols <- cluster_cols[cluster_cols %in% clusters_of_interest]
de_genes %>% 
  filter(p_val_adj < 0.05 & abs(avg_log2FC) > 0.3 & cpdb.secreted) %>%
  select(X) -> sasp
sasp <- sasp$X
names(sasp) <- rep("#00FFFF", length(sasp))
sasp <- sasp[!duplicated(sasp)]

net <- dom_df_to_igraph_lig_rec(out.net.sub.stromal, send_cols = send_cols, rec_cols = rec_cols, label_additional = sasp)
coords <- net_layout_lig_rec(net, out.net.sub.stromal)
plot_network(net, coords, 0)
```

# save plot and data table
```{r, echo=FALSE, eval=FALSE}
svglite::svglite(here::here("plots_for_figures", "plots", "Fig_3", "Endo_vascular_to_stromal.no_collagens_only_nonSnC_recv.svg"), height = 10)
plot_network(net, coords, 0)
dev.off()

write.csv(out.net.sub.stromal, here::here("plots_for_figures", "data_tables", "Fig_3", "data_in_plots", "SnC_vascular_to_stromal_vascular.csv"))
```


# Chord diagram

# Chord plot function
```{r}
circos_plot <- function(network_slice) {
  set.seed(1)
  chordDiagram(network_slice[,c("receptor", "transcription_factor", "cluster_cor")],
              col = network_slice$col, 
              big.gap = 20,
               #scale = TRUE,
              directional = 1, 
              direction.type = c("diffHeight", "arrows"),
              link.arr.type = "big.arrow", 
              annotationTrack = "grid",
              preAllocateTracks = list(list(track.height = 0.2),
                                       list(track.height = 0.08)))

  circos.track(track.index = 1,
               panel.fun = function(x, y) {
               circos.text(CELL_META$xcenter, 
                           CELL_META$ylim[1], 
                           #str_wrap(CELL_META$sector.index, width = 20), 
                           str_split(CELL_META$sector.index, pattern = "\\.") |> unlist() |> tail(1),
                           facing = "clockwise", 
                           cex = 0.8, 
                           adj = c(0, 0.5), 
                           niceFacing = TRUE)},
               bg.border = NA)
circos.clear()
}
```


```{r}
rec <- c("TEK", "NOTCH3", "NOTCH4", "PDGFRA", "PDGFRB", "TGFbeta_receptor2", "RAMP3", "TNFRSF1A")
tf <- c("HIF1A", "ERG", "ETS1", "KLF9", "PBX3", "BCL6B", "TBX21", "ZNF629", "SOX18", "SMAD1", "RARA", "E2F1", "LTF", "TEAD4", "SRF", "RFX2", "ETS2", "NR3C1", "GATA6", "THRA", "ALX1", "ELK3")

cols <- setNames(names(cluster_cols), cluster_cols)
network_slice <- out.net.sub.stromal %>%
  filter(receptor %in% rec & transcription_factor %in% tf) %>%
  mutate(col = recode(receiving_cluster, !!!cols)) %>% 
  select(receptor, transcription_factor, cluster_cor, receiving_cluster, col) %>%
  filter(cluster_cor > 0.1) %>%
  distinct()

# Cluster legend
network_slice %>% 
  select(receiving_cluster, col) %>%
  distinct %>%
  deframe -> cl_cols
cl_legend = Legend(labels = names(cl_cols), title = "Cluster", legend_gp = gpar(fill = cl_cols))

circos_plot(network_slice)
draw(cl_legend, x = unit(2, "mm"), y = unit(6, "mm"), just = c("left", "bottom"))
```

# save chord diagram and data table
```{r, echo=FALSE, eval=FALSE}
pdf(here::here("plots_for_figures", "plots", "Fig_3", "Endo_vascular_to_stromal.receptor_TF.chord_exploratory.pdf"))
circos_plot(network_slice)
draw(cl_legend, x = unit(2, "mm"), y = unit(6, "mm"), just = c("left", "bottom"))
dev.off()

write.csv(network_slice, here::here("plots_for_figures", "data_tables", "Fig_3", "data_in_plots", "vascular_stromal_Rec_TF_correlations_downstream_of_SnC_signals.csv"))
```


```{r}
sessionInfo()
```









